---
title: "PROJECT TITLE"
author: "Duke Squirrels"
date: "03/22/18"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-packages, message=FALSE}
library(tidyverse)
```


## Section 1. Introduction

The data that we obtained contains information regarding famous figures throughout history. The data came from https://www.kaggle.com/mit/pantheon-project/data which originated from The Massachusetts Institute of Technology about a year ago. The data is based off of wikipedia pages. We felt that this data is credible given that it originates from an institution like MIT. The variables within the data are as follows: `full_name`, `sex`, `birth_year`, `city `, `state`, `country`, `continent`, `latitude`, `longitude`, `occupation`, `industry`, `domain`, `article_langauges`, `page_views`, `average_views`, and `historical_popularity_index`. Given the information, all of us had many ideas as to questions to figure out. 

## Section 2. Data analysis plan

## Section 3. Data

```{r load-data, warning=FALSE, message=FALSE}
panth <- read_csv("data/database.csv")
```

