---
title: "Wikipedia Popularity"
author: "Duke Squirrels"
date: "04/19/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Load Packages

```{r load-packages, message=FALSE}
library(tidyverse)
library(broom)
```

## Load Data

```{r load-data, message=FALSE}
panth <- read_csv("data/cdatabase.csv",
                  na = c("Unknown", "1237?", "530s"))
```

## Introduction
 
The data that we obtained contains information regarding historical figures. We downloaded the data from Kaggle, but the data was collected by the Massachusetts Institute of Technology about a year ago. The data is based off of metrics from many wikipedia pages and believe the variables in the dataframe can be used to extrapolate what makes a historical figure "popular" by Wikipedia standards.
 
By the end of our data analysis, we aim to derive the perfect combination of variables that lead to a high popularity index, which is recorded in the dataframe.
 
### Section 1- Introduction to the Data

```{r filter_NAs_rank_data}
panth_mod <- na.omit(panth)

panth_mod <- panth_mod %>%
  mutate(rank = 1:nrow(panth_mod))

panth_mod %>%
  select(rank, full_name, occupation, birth_year, historical_popularity_index) %>%
  head(10)
```
 
There are `r ncol(panth_mod)` variables and `r nrow(panth_mod) ` observations (with all NAs removed in the new dataframe). Before removing the NAs, the full dataframe had 11,341 observations.
 
In addition, we decided to add a `rank` variable to the dataframe to rank the historical figures based off of their `historical_popularity_index`.

We have also found the top 10 figures with the highest popularity index score in order. They are:
 Aristotle, Plato, Jesus Christ, Socrates, Alexander the Great, Leonardo da Vinci, Julius Caesar, Homer, Pythagoras, Archimedes

#### Introductory Exploratory Analysis

```{r counting_continents} 
panth_mod %>%
  count(continent) %>%
  arrange(desc(n))
```

Europe  has the most historical figures in the dataset with 6073, followed by North America with 2351, and then by Asia, which has 1021.

```{r page_views_visual}
panth_mod %>%
  ggplot(mapping = aes(x = continent, y = average_views)) +
  geom_boxplot() + 
  coord_flip() +
  labs(title = "Distribution of Average Page Views", y = "Average Page Views", 
       x = "Continent", subtitle = "by Continent")
```

```{r outlier_avg_page}
panth_mod %>%
  group_by(continent) %>%
  select(rank, full_name, country, continent, average_views,industry,occupation)%>%
  arrange(desc(average_views))%>%
  top_n(n = 1, wt = average_views)

```

Visualized here is a distribution of the average number of page views, which is defined as the the average number of page views across all of the different language Wikipedia articles of the same figure, of historical figures by continent. Each point in this visual represents a historical figure's average number of Wikipedia page views. It is clear that North America has many individuals whose articles have a high average number of page views. Furthermore, the top figures who lead their continent in average page views tend to be entertainers of sorts.

```{r prop_top1000}
panth_mod %>%
  top_n(n = 1000, wt = historical_popularity_index)%>%
  count(continent)%>%
  mutate(prop = n/sum(n))
```
Even though North America had many individuals with high average number of page views, they only contribute to 9% of the top 1000 figures ranked by historical popularity index. This tells us that historical popularity index likely does not weigh average page views as highly as other variables. A high average number of page views does not mean a high ranking in terms of historical popularity. As evidenced by the top average view earners such as Kim Kardashian or Ronaldo, who have low ranks in historical popularity.

### The Historical Popularity Index

According to the official codebook from kaggle, the historical popularity index value measures approximately how popular the article on the historical figure was and how well known the figure was. It was created by "adding information on the age of the historical character, the concentration of page views among different languages, the coefficient of variation in page views, and the number of page views in languages other than English."

(https://www.kaggle.com/mit/pantheon-project)

```{r popularity-across-globe, fig.width=13, fig.height=5}
ggplot(panth_mod, mapping = aes(x = longitude, y = latitude)) +
  geom_point(aes(color = historical_popularity_index)) +  scale_colour_gradient2() +
  theme_minimal() +
  labs(title = "Historical Popularity Index Around the World", x = "Longitude", 
       y = "Latitude", color = "Popularity Index Gradient")
```
 
From this visual, we can see that the areas of the world that are generally uninhabitable or extremely rural do not have historical figures. For example, the center of South America, a large portion of North Africa, most of Russia, and a big portion of Australia lack a concentration of historical figures. Additionally, in these countries and continents, the coast seems to have the most historical figures. This map shows a gradient measuring the historical popularity index around the world, with the lightest points representing figures with the highest popularity index scores.
 
For the first part of our analysis, we will look at the variables `domain`, `sex`, and `article_languages` indpendently using simple linear regression and later create a full linear model to determine which variables are most important when calculating the popularity index score of historical figures across time.

### Section 2 - How the Variable `Domain` Affects `Historical_Popularity_Index`
```{r distribution-of-index}
panth_mod %>%
  ggplot(mapping = aes(x = historical_popularity_index)) +
  geom_histogram(bins = 30) + 
  facet_wrap(~domain) +
  labs(title = "Distribution of Historical Popularity Index Scores", y = "Count", 
       x = "Historical Popularity Index Score", subtitle = "by Domain")

panth_mod %>%
  group_by(domain)%>%
  summarise(mean = mean(historical_popularity_index), 
            median = median(historical_popularity_index))
```

```{r looking_at_data}
panth_mod %>%
  count(domain) %>%
  arrange(desc(n))
```

The Arts and Institutions have the most historical figures in the dataset, with 2767 and 2753, respectively. Exploration and Business & Law had the least number of people with 103 and 88. Looking at the summary statistics, all domains except Sports had similar means and medians (historical popularity index), which were near 22, while Sports had values near 17. 

#### Simple Linear Regression

```{r domain_linear_model}
domain_m <- lm(historical_popularity_index ~ domain, data = panth_mod)
tidy(domain_m) %>%
  select(term, estimate)
```

Here we estimated the historical popularity index using the `domain` variable using a simple linear regression. The slope for the level in `domain` named `Business & Law` is 0.502, suggesting that historical figures who belong in the `Business & Law` domain have, on average, an increase in their overall popularity index of 0.502 as long as all other variables are held constant.

In this model, the largest coefficient belongs to the domain of `Humanities`. With all else held constant, historical figures who belong to the domain of `Humanities` have a historical popularity index that is, on average, 2.448 higher than the intercept. 

The least coefficient belongs to `Sports`. With all else held constant, historical figures who belong to `Sports`, have an index that is on average, 4.11 lower than the intercept. 
 
The linear model, based on the output, is:
 
`(historical_popularity_index) = 21.839(intercept) + 0.502(domainBusiness & Law) + 1.710(domainExploration) + 2.448(domainHumanities) + 1.716(domainInstitutions) + 0.532(domainPublic Figure) + 1.534(Science & Technology) - 4.112(domainSports)`

```{r r-squared_domain}
glance(domain_m)$r.squared
```

We found that the r-squared for the linear model `domain_m` is 40.1%, which suggests that 40.1% of the variability of the data can be explained by the domain of the historical figure. 
 
### Section 3 - How the Variable `Sex` Affects `Historical_Popularity_Index`

```{r men-and-women}
panth_mod %>%
  count(sex) %>%
  mutate(prop = n/sum(n))
```

Based on the filtered dataframe, there are 1,427 women and 8,852 men that are considered historical figures of the total 10,279 historical figures. There are about 6.2 times as many historical men than women overall in the data. The timeframe of this data starts at `r panth_mod%>%arrange(birth_year)%>%select(birth_year)%>%head(1) `, or 3500 BCE, and ends at `r panth_mod%>%arrange(desc(birth_year))%>%select(birth_year)%>%head(1)`, spanning about 5000 years. This means that a mere 13.9% of women in the entire timeframe are considered historical figures.

#### Simple Linear Regression

```{r estimating_popularity_sex}
m_pop_sex <- lm(historical_popularity_index ~ sex , data = panth_mod)
tidy(m_pop_sex) %>%
  select(term, estimate)
```

Here we estimated the historical popularity index using the `sex` variable using a simple linear regression. The slope for the categorical variable `sexMale` is 1.55, suggesting that historical figures who are men have, on average, an increase in their overall popularity index of 1.55 as long as all other variables are held constant. 
 
The linear model, based on the output, is:
 
`(historical_popularity_index) = 20.8(intercept) + 1.55(sexMale)`
 
```{r r-squared_sex}
glance(m_pop_sex)$r.squared
```

We found that the r-squared for the linear model `m_pop` is 2.54%, which suggests that 2.54% of the variability of the data can be explained by sex and that the model does not fit our data very well. We think there might be a confounding variable, so we will look to see if `birth_year` is one.

#### Simple Linear Regression - For All Figures Born After 1920

```{r women-after-1920}
panth_mod%>%
  filter(birth_year > 1920)%>%
  count(sex)%>%
  mutate(prop = n/sum(n))
```

For historical figures born after 1920, there are about 4 times as many male historical figures than female figures. This is intriguing because in this 85 year timeframe from 1920 - 2005, we have 1053 historical women out of a total of 1427 women in the entire timeframe. 

```{r new_m_pop_sex}
m_pop_sex2 <- lm(historical_popularity_index ~ sex , data = panth_mod %>%
                   filter(birth_year > 1920))
tidy(m_pop_sex2) %>%
  select(term, estimate)
```

In the previous model, we predicted the `historical_popularity_index` by `sex` across the entire ~5000 year time period of the data. The result was that historical figures who were men had, on average and with all other variables held constant, a popularity index score that was 1.55 points higher than that of women who were historical figures. However, in this model, we thought it would be interesting to analyze the 85 year timeframe after the year 1920, when women were given the right to vote in the U.S. and when, later in the century, women across the world where also granted greater rights. As a result, women made up about 20% of the historical figures as opposed to making up 13.9% of the historical figure population in the previous analysis.
 
The resulting linear model that only looked at the historical figures after 1920 is as follows:
 
`(historical_popularity_index) = 19.6(intercept) + 0.569(sexMale)`
 
The slope of the `sexMale` variable decreased significantly from the previous analysis. This shows that time is a factor that affects the historical popularity index of women specifically.

```{r r-squared_after_1920}
glance(m_pop_sex2)$r.squared
```

We found that the r-squared for the linear model `m_pop_sex2` is 0.51%, which suggests that 0.51% of the variability of the data can be explained by the linear model and that the model does not fit our data very well at all. The fact that the r-squared value went down makes sense because it is harder to differentiate between men and women, in terms of their historical popularity index score. But in larger picture, we see that sex as a predictor does not impact the variability to a large extent and there are other variables like domain which are more impactful. 

### Section 4 - How the Variable `Article_Languages` Affects `Historical_Popularity_Index`

#### Simple Linear Regression

```{r article_languages_linear_model}
artlang_m <- lm(historical_popularity_index ~ article_languages, data = panth_mod)
tidy(artlang_m) %>%
  select(term, estimate)
```

Here we estimated the historical popularity index using the `article_langugaes` variable on a simple linear regression model. The slope for the variable `article_languages` is 0.089, suggesting that for every one increase in the amount of languages the article has been translated into, that historical figure, on average, will have an increase in their overall popularity index of 0.089 as long as all other variables are held constant.
 
The linear model, based on the output, is:
 
`(historical_popularity_index) = 21.839(intercept) + 0.089(article_languages)`

```{r r-squared_article_languages}
glance(artlang_m)$r.squared
```

We found that the r-squared for the linear model `artlang_m` is 21.6%, which suggests that 21.6% of the variability of the data can be explained by the linear model and that the model doesn't necessarily fit our data very well, but it fits it significantly better than the sex variable.

```{r visualizing-index-by-popularity}
panth_mod %>%
  filter(article_languages > 75 & continent != "Oceania")%>%
  ggplot(mapping = aes(y = historical_popularity_index, x = article_languages,
                       color = factor(continent))) + geom_point(alpha = .5) + 
    geom_smooth(method = "lm", se = FALSE) +
    labs(x = "Number of Article Languages", y = "Popularity Index", 
    title =
      "Effect of # of languages the article is translated into on popularity index", 
    color = "Continent") 
```

From this visual, for historical figures from Asia and Europe, we see that as the number of languages one's article increases, the popularity index also increases. However, for Africa and the Americas, there is little upward movement in popularity as article languages increases. The regression lines for these three continents tend to stay relatively constant.

```{r greater-than-200}
panth_mod %>%
  filter(continent == "Asia")%>%
  select(full_name, historical_popularity_index, article_languages, country,occupation) %>%
  arrange(desc(article_languages))%>%
  head(3)
```

There is a historical figure that is an outlier considering Asia's spread in the visual. By filtering the data for the figure with a page translated into more than 200 languages, we identified this individual as Jesus Christ, whose Wikipedia page has been translated into 214 different languages. The next highest was Muhammed, whose article has 150 languages. Both were historical figures.

```{r who-is-the-outlier}
panth_mod %>%
  select(full_name, historical_popularity_index, article_languages, country) %>%
  filter(historical_popularity_index < 20, article_languages > 175)
```

Interestingly, one of the historical figures has a low popularity index score, an outlier, but his/her wikipedia page has been translated into almost 200 different languages. After filtering the data to locate the point that had a popularity index score less than 20 and an article that was translated into more than 175 languages, we derived that the outlier historical figure from North America is Corbin Bleu, the actor from High School Musical.
 
After looking at how the variables `domain`, `sex`, and `article_languages` each affect historical popularity index, we see that they do play a role. To summarize our results and check to see if other variables affect the historical popularity index all at once, we will use a multiple linear regression full model.
 
#### Multiple Linear Regression

```{r full_model}
m_full <- lm(historical_popularity_index ~ sex +
               domain +
               article_languages +
               continent +
               birth_year,
             data = panth_mod)
tidy(m_full) %>%
  select(term, estimate)
```

Here we estimated the historical popularity index using the `sex`, `domain`, `birth_year`, `article_languages`, and `continent` variables. We also included the interaction between continent and article languages. We would interpret the slope the same way we did with the simple linear regression above that had the `sex` variable only.

The linear model, based on the output, is:

`(historical_popularity_index) = 20.8(intercept) + 1.43(sexMale) + 0.335(domainBusiness & Law) + 0.290	(domainExploration) + 1.099(domainHumanities) + 0.394(domainInstitutions) + 0.753(domainPublic Figure) + 0.781(domainScience & Technology) - 4.20(domainSports) + 0.0675(article_languages)	+ 0.290	(continentAsia) + 1.224(continentEurope) + 0.411(continentNorth America	) - 0.204(continentOceania) + 1.155(continentSouth America) - 0.00182(birth_year)`

```{r full_r-squared}
glance(m_full)$r.squared

glance(m_full)$adj.r.squared
```

The r-squared value for the full model is 63.5%, which means that more than half of the variability of the data can be explained by the model. This is a significant increase from the sex regression model we did comparing `sex` and `historical_popularity_index` scores, which resulted in an r-squared of 2.54%. The adjusted r-squared is 63.4%. This is also a large increase from the domain model, which had an r-squared of 40%. Thus the full model has the most predictive power. 

#### Backwards Selection with AIC

```{r backwards_selection}
selected_model <- step(m_full, direction = "backward")

tidy(selected_model) %>%
  select(term, estimate)

glance(selected_model)$AIC
glance(selected_model)$r.squared
glance(selected_model)$adj.r.squared
```

After creating the selected model, we found that the full and selected models were identical, which indicates that the full model had the best predictive power. 

#### The perfect historical popularity index

Based on the full and selected models, to have the highest popularity index score, one should:
be a man, study in the domain of the humanities, and live somewhere in the continent of South America. Additionally, the predicted popularity index score would increase if the figure was born before the year 0 (or before common era). This is because though the slope for `birth_year` is negative, the birth year itself for these figures is also negative (ie -3500), so the overall slope would be positive. 

```{r perfect_index_score}
panth_mod %>%
  filter(continent == "Europe",
         birth_year < 0,
         sex == "Male",
         domain == "Humanities") %>%
    select(rank, full_name, historical_popularity_index, occupation)
```

After filtering for the characteristics deemed most impactful on historical popularity score, we see that it does indeed result in very high ranking individuals. The top 5 in this list are all from the top 10 of the original ranking by popularity index.

## Conclusion
When we found the data, we thought it would be interesting to analyze what makes a person popular on Wikipedia by MIT’s standards, or rather how the historical figures held up against each other based on MIT’s calculated Historical Popularity index. After thorough analysis that began with visualizing popularity index scores across the world and calculating summary statistics based on the figures’ average page views, we started to find that there were inconsistencies among the historical figures’ popularity scores and their page views. For example, figures with the highest number of average page views across the continents are entertainers with mediocre popularity scores; these are historical figures like Chris Hemsworth, Kim Kardashian, and Lionel Messi. Additionally, the top 10 figures with the highest popularity index scores were all European men who studied the humanities, with the exception of Archimedes, who studied math, and Alexander the Great, who was a military figure. From analyzing the interactions between the variables of this dataset, we have gathered that the world is very possibly more interested in figures who have shaped our society than figures who seek to entertain.To answer our research question, which focused on finding the variables/characteristics that maximized a historical figure’s popularity index score, we created a full model and a selected model, though the selected model, created using the AIC, kept all of the full model’s variables, suggesting that our full model had the strongest predictive power.  We concluded that a man from Europe who studied the humanities and lived before the common era would have the highest popularity score. When we filtered our dataframe for these variables, many of the resulting figures were already in the list of top 10 figures with the highest popularity index scores. 

As for the reliability and validity of the data, we find that because we used an already-calculated score for our main analysis (popularity index), we were subjected to MIT’s calculations of the scores and their interpretations of the variables that went into calculating the historical popularity index variable. To improve our data analysis, we could have used the entire dataframe instead of using a filtered one that removed ~1000 observations. This could have resulted in some greater change in the model results and ultimately changed our entire analysis. 

If we were able to start the project over, we would choose to web scrape our own data in order to have the data we specifically needed to conduct the analyses we wanted to. For example, we wanted to conduct hypothesis testing on our dataset, but came to the realization that our dataframe represented a population rather than a random sample of historical figures, so any of the findings of our test couldn’t be applied to the general population because we already had the population with us. Furthermore, if we had had more time and more knowledge, we would have liked to create an outline of the world map and plot the points of our dataframe onto that outline to get a better visual of the world. Additionally, with more information and some help, we would have liked to continue the project by conducting timeline analyses to observe how the passage of time affected the historical popularity index scores among the different domains. 


